\documentclass{tufte-handout}

\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{graphicx}
\setkeys{Gin}{width=\linewidth,totalheight=\textheight,keepaspectratio}
\graphicspath{{./graphics/}} % DO RELATIVE FILE PATHS EVEN WORK??? 

\title{Block Practice — Round 1}
\author[IA Math Team]{}
\date{24 September 2025}

\usepackage{booktabs}
\usepackage{units}
\usepackage{fancyvrb}
\fvset{fontsize=\normalsize}

\usepackage{multicol}

\usepackage{lipsum}

\newcommand{\doccmd}[1]{\texttt{\textbackslash#1}}
\newcommand{\docopt}[1]{\ensuremath{\langle}\textrm{\textit{#1}}\ensuremath{\rangle}}
\newcommand{\docarg}[1]{\textrm{\textit{#1}}}
\newenvironment{docspec}{\begin{quote}\noindent}{\end{quote}}
\newcommand{\docenv}[1]{\textsf{#1}}
\newcommand{\docpkg}[1]{\texttt{#1}}
\newcommand{\doccls}[1]{\texttt{#1}}
\newcommand{\docclsopt}[1]{\texttt{#1}}

\begin{document}

\maketitle

\begin{abstract}
\begin{itemize}
\item This is the first round of a ``practice'' ``mock'' tournament.
\item Please make sure you are working in a team. Despite working in a team, please be mindful of loudness: some teams might be working on the exploration round.
\item You may submit your answers for verification; if every answer is correct, you will move on to the next round.
\item Please note the round progression on the board to potentially develop a strategy.
\end{itemize}
\end{abstract}

\vspace{1.5em}

\begin{enumerate}
    \setlength{\itemsep}{10em}
    \item Given $x-y=4$ and $x+y=10$, find the quantity $x^2+y^2$.
    \item The sum of five consecutive odd integers is $-2025$. What is the last number?
    \item A coin is flipped exactly three times. What is the probability that heads comes up exactly twice?
    \item Sreyas Sabbani the Spaghetti Slicer Slices Six Slices of Spaghetti in Seven Seconds. Sreyas Slices $s$ Sections in Seventy Seconds.
    
    Solve $s$.

    
\end{enumerate}

\end{document}
